---
title: "Report_Leader_DC_reanalysis"
author: "Abdenour ABBAS "
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: show
theme: lumen
editor_options: 
  chunk_output_type: inline
---

```{r setup, message=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center",
               fig.retina = 2,
               fig.width = 10,
               cache = TRUE,
               cache.lazy = FALSE,
               # max.print = 10,
               out.lines = 10)

# Load libraries
library(Seurat)
library(ggplot2)
library(dplyr)
library(magrittr)
library(parallel)
library(tidyverse)
library(qs)
library(presto)
library(simspec)
library(nichenetr)
library(patchwork)
library(nichenetr,quietly = T)

# Load utility scripts
config <- yaml::read_yaml("/pasteur/helix/projects/Dcai/Abdenour/analysis/utils/config.yaml")
source(paste0(config$path_to_scripts, "plot_heatmap2.R"))
source(paste0(config$path_to_scripts, "utils.R"))
source(paste0(config$path_to_scripts, "genes_related.R"))
# Force Seurat to use v3 assays 
options(Seurat.object.assay.version = 'v3') # https://github.com/satijalab/seurat/issues/8164

# Use custom colors in ggplot2
options(ggplot2.discrete.colour=pal)
options(ggplot2.discrete.fill=pal)
# scales::show_col(pal)
```

# 1. Wound Healing Dataset
```{r Load Data, eval=F}
# ---- PARAMETERS ----
# Path where all 10X subfolders are located
data_dir <- "../../../../../data_global/public/single_cellRNA/Skin/GSE241132_human skin wound healing/GSE241132_RAW/"  # <- change this

# Get all subfolders (each one should contain matrix.mtx.gz, barcodes.tsv.gz, features.tsv.gz)
folders <- list.dirs(data_dir, full.names = TRUE, recursive = FALSE)

# ---- LOAD ALL DATASETS ----
seurat_list <- list()

for (folder in folders) {
  sample_name <- basename(folder)  # use folder name as sample ID
  cat("Loading:", sample_name, "\n")
  
  counts <- Read10X(data.dir = folder)
  seurat_obj <- CreateSeuratObject(counts = counts, project = sample_name)
  
  # Store sample ID in metadata
  seurat_obj$sample <- sample_name
  
  seurat_list[[sample_name]] <- seurat_obj
}

# ---- MERGE INTO ONE OBJECT ----
# Take the first as reference, then merge all others
for (i in names(seurat_list)) {
  seurat_list[[i]] <- RenameCells(seurat_list[[i]],
                                         add.cell.id = i)
}

merged <- seurat_list[[1]]
if (length(seurat_list) > 1) {
  for (i in 2:length(seurat_list)) {
    merged <- merge(merged, y = seurat_list[[i]])
  }
}

# extract the D + number part
merged$timepoint =  sub(".*(D[0-9]+)$", "\\1", merged$orig.ident)
merged$timepoint = factor(merged$timepoint, levels=c("D0","D1","D7","D30"))
merged$patient_ID =  sub("(PWH[0-9]+).*$", "\\1", merged$orig.ident) 
merged$barcode = colnames(merged)
merged = PercentageFeatureSet(merged, pattern = "^MT-", col.name = "percent.mt")
# ---- SAVE RESULT ----
VlnPlot(merged, "nFeature_RNA")+ scale_y_log10()
# cat("✅ Done! Saved merged Seurat object as merged_seurat.rds\n")
```

```{r eval=F}
merged = qs::qread("../objects/human_skin_wound_healing_GSE241132.qs")
```

## 1.1. Subsetting myloid cells
```{r eval=F}
# 65521  cells
merged %<>% subset(subset= nFeature_RNA>500) %<>% subset(percent.mt <20) 
# 62134 cells
# plot(density(merged$percent.mt)) 
# plot(density(metadata$nCount_RNA))
# table(merged$nFeature_RNA<501) 
metadata <- read.delim("../../../../../data_global/public/single_cellRNA/Skin/GSE241132_human skin wound healing/GSE241132_cell_metadata.txt")
merged$cell_type = metadata$newMainCellTypes[match(merged$barcode, metadata$barcode)]
merged$subtype = metadata$newCellTypes[match(merged$barcode, metadata$barcode)]
DimPlot(merged, group.by = "subtype")
# seu_sub = subset(merged, cell_type=="Myeloid", invert=F)
# qsave(seu_sub, "human_skin_wound_healing_GSE241132_Myeloid_filtered.qs", nthreads = 8)
```

## 1.1. Subsetting myloid cells
```{r eval=T}
seu_sub = qs::qread("human_skin_wound_healing_GSE241132_Myeloid_filtered.qs", nthreads = 8)
```

Let's exclude some gene lists from the clustering / visualization as they may hide relevant biological signals.
```{r Prepare gene masks, eval=T}
all_genes <- rownames(seu_sub)
HB_genes <- c("HBB","HBA1","HBA2", "HBD","HBE1","HBG1","HBM","HBBP1","HBQ1","AHSP")
metallothionein=c('MT1HL1','MT2A','MT1E','MT1M','MT1A','MT1B','MT1F','MT1G','MT1H','MT1X')
rpgenes= grep("^RP[LS0-9]", all_genes, value=T) # 
other=c("MALAT1","XIST","JCHAIN")
hsp_genes=grep("^HSP", all_genes, value=T) 
dnaj_genes=grep("^DNAJ", all_genes, value=T)
stress_genes= c("JUN","FOS","FOSB","JUNB","ATF3","ATF4",dnaj_genes,hsp_genes)
mtgenes=grep("^MT-", all_genes, value=T) 
MTRN_genes=grep("^MTRN", all_genes, value=T)
IGgenes = grep("^IG[KLH]", all_genes, value=T)
HLAgenes = grep("^HLA-", all_genes, value=T)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
cell_cycle_genes = c(s.genes,g2m.genes)
genes_mask_clust = c(cell_cycle_genes,HB_genes, metallothionein, rpgenes,other,IGgenes,HLAgenes,MTRN_genes,mtgenes) #stress_genes
genes_mask_plot = c(HB_genes, metallothionein, rpgenes,other,IGgenes,MTRN_genes,mtgenes)#stress_genes
```

## 1.2. DimReduction 
```{r }
DefaultAssay(seu_sub) = "RNA"
seu_sub = NormalizeData(seu_sub)
seu_sub <- CellCycleScoring(seu_sub, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
seu_sub = FindVariableFeatures(seu_sub)
vargenes = setdiff(VariableFeatures(seu_sub), genes_mask_clust)
seu_sub <- ScaleData(seu_sub, assay = "RNA", features = vargenes, vars.to.regress =  c("S.Score", "G2M.Score"))
seu_sub <- RunPCA(seu_sub,verbose = FALSE, npcs = 50, reduction.name = "pca", features = vargenes)
nPC= 15
ElbowPlot(seu_sub, ndims = 50, reduction = "pca")+geom_vline(xintercept=nPC,  color="red")
seu_sub <- RunUMAP(seu_sub, reduction = "pca",min.dist = 0.5,spread = 1,  dims = 1:nPC, reduction.name = "umap")
```

## 1.3. Clustering

```{r fig.height=9, fig.width=12}
DimPlot(seu_sub)

# 36600 features across 65521
seu_sub <- FindNeighbors(seu_sub, reduction = "pca", dims = 1:nPC)
seu_sub <- FindClusters(seu_sub, resolution = 0.2)
# Cluster 6 are DC, subset object to cluster 6 and recluster
DimPlot(seu_sub)
 
cells= Cells(seu_sub)
FeaturePlot(seu_sub, features=c("FCER1A","CD1C","C5AR1","FOLR2","CD207","SPP1","SELENOP","TREM2"), cells = cells, ncol=4, reduction= "umap") & NoAxes()&coord_equal()
markers = presto::wilcoxauc(seu_sub)
markers_dist = markers %>% filter(!feature %in% genes_mask_clust) %>% arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>%  group_by(group) %>% arrange(desc(auc)) %>% slice_head(n = 30)
```

## 1.4. Annotation and reordering of clusters
```{r eval=F}
VlnPlot(seu_sub, features =c("CD1C","FCER1A","CLEC10A","CD207","CD1A","EPCAM","CCR6","VCAN","FCN1","C5AR1","IDO1","CCR7","CLEC9A","CD14"), fill="ident",stack=T, flip=T, group.by = "annot_lev3")+NoLegend()+plot_layout(ncol=3)
p <- SCpubr::do_CorrelationPlot(sample = seu_sub,
                                group.by = "RNA_snn_res.0.2",
                                legend.position = "right")
# levels(p$data$x)
seu_sub$seurat_clusters = factor(seu_sub$RNA_snn_res.0.2, levels= c("4","1","2","0","3","5","6","7"))  # levels(p$data$x)

seu_sub$res.0.2 = plyr::mapvalues(seu_sub$seurat_clusters, from = levels(seu_sub$seurat_clusters), to =c(1:8))
seu_sub$annot_lev3 = plyr::mapvalues(seu_sub$res.0.2, from = c(1:8), to =c("MonoMac_1","MonoMac_2","MonoMac_3","DC2_CCR7","DC2_CD1A_CCR7","DC1","mature_DC","LC"))
seu_sub@active.ident = seu_sub$annot_lev3
```


## 1.4. Visualization
Heatmaps
```{r , fig.height=10.5, fig.width=11.5, eval=T}
seu_sub$seurat_clusters = seu_sub@active.ident
markers = presto::wilcoxauc(seu_sub, group_by="seurat_clusters")
top_markers= presto_topn_auc(markers,object= seu_sub@active.ident,min_pct_in=10, genes_filter=genes_mask_plot, top_n = 15,minavgExpr=0.1)

seu_sub <- ScaleData(seu_sub,features=c(top_markers$feature))


plot_heatmap2(seu_sub, markers =unique(c(top_markers$feature)),sort_var = c("seurat_clusters","subtype","timepoint"),
              anno_var = c("seurat_clusters","subtype","timepoint"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,T,T,T,F),padding = unit(c(0, 0, 0,2), "cm"),
              row_font_size = 7)

VlnPlot(seu_sub, features =c("CD1C","FCER1A","CLEC10A","CD207","CD1A","EPCAM","CCR6","ITGAE","HPGDS","IL18R1","SIGLEC9","CD72","LAMP3","IDO1","CCR7","CLEC9A","CD14"), fill="ident",stack=T, flip=T)+NoLegend()+plot_layout(ncol=3)

top_markers= presto_topn_auc(markers,object= seu_sub@active.ident,min_pct_in=10, max_pct_out = 100, genes_filter=genes_mask_clust, top_n = 100,minavgExpr=0.1)

seu_sub <- ScaleData(seu_sub,features=c(top_markers$feature))

plot_heatmap2(seu_sub,  markers = unique(c(top_markers$feature[top_markers$group=="7"])),sort_var = c("seurat_clusters","subtype","timepoint"),
              anno_var = c("seurat_clusters","subtype","timepoint"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,T,T,T,F),padding = unit(c(0, 0, 0, 5), "cm"),
              row_font_size = 7)

markers = presto::wilcoxauc(seu_sub) 

markers$feature = UpdateSymbolsHuman(markers$feature)

top_markers = markers %>% filter(!feature %in% genes_mask_plot & avgExpr>0.1 & logFC>0.1  & pct_in>10) %>% group_by(group) %>% arrange(desc(auc),.by_group =T )

top_markers_distinct = markers %>% filter(!feature %in% genes_mask_plot &  avgExpr>0.1 & pct_in>10 & logFC>0.1)  %>% arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>% group_by(group) %>% arrange(desc(auc), .by_group =T)

# WriteXLS::WriteXLS(top_markers_distinct, "results/skin_wound_healing_myeloid_clusters_markers_distinct.xlsx")
# WriteXLS::WriteXLS(top_markers, "results/skin_wound_healing_myeloid_clusters_markers.xlsx")
```



## 1.5. Saving Data
```{r}
# qsave(seu_sub, "human_skin_wound_healing_GSE241132_Myeloid_filtered_processed.qs", nthreads = 8)
seu_sub = qs::qread("human_skin_wound_healing_GSE241132_Myeloid_filtered_processed.qs", nthreads = 8)
```

 ################################################################################
 ################################################################################
 ################################################################################

# 2. Multi-site Skin Atlas
Multi site Skin Atlas (biorXiv)
```{r}
# atlas_skin_seu = readRDS("../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3.RDS")
atlas_skin_seu@assays$RNA$scale.data = NULL
# qsave(atlas_skin_seu, "../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3.qs", nthreads = 8)
atlas_skin_seu = qread("../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3.qs", nthreads = 8)
```
## 2.1 Subsetting myeloid cells

```{r}
atlas_skin_seu= NormalizeData(atlas_skin_seu)
DimPlot(atlas_skin_seu)
str(atlas_skin_seu)
atlas_skin_myelo = subset(atlas_skin_seu, celltype=="Myeloid Cells") 
# qsave(atlas_skin_myelo, "../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3_myelo.qs", nthreads = 8)
atlas_skin_myelo = qread("../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3_myelo.qs", nthreads = 8)
str(atlas_skin_myelo)
```

```{r}
all_genes <- rownames(atlas_skin_myelo)

HB_genes <- c("HBB","HBA1","HBA2", "HBD","HBE1","HBG1","HBM","HBBP1","HBQ1","AHSP")
metallothionein=c('MT1HL1','MT2A','MT1E','MT1M','MT1A','MT1B','MT1F','MT1G','MT1H','MT1X')
rpgenes= grep("^RP[LS0-9]", all_genes, value=T) # 
other=c("MALAT1","XIST","JCHAIN",grep("^IGLL", all_genes, value=T))
mtgenes=grep("^MT-", all_genes, value=T) 
MTRN_genes=grep("^MTRN", all_genes, value=T)
# stress_genes = c("HSPA1B","HSPA6","DNAJB1","HSPA1A","HSPB1","JUN","UBB","DNAJB4","BAG3","UBC","SOD1","RGS2","EIF4A2","SERPINH1","DEDD2","CLK1","AHSA1","COMMD6","HSP90AA1","HSPH1","HSPE1","HSPD1","HSP90AB1","DUSP1","HSPA8","FOS","DNAJA1","PPP1R15A","CACYBP","KLF6","ZFAND2A","NR4A1","FOSB","DNAJA4","CHORDC1","MRPL18","BTG2","FKBP4","NEU1","ATF3","MYADM","ERN1","DOK2","GADD45B","TXNIP","NABP1","SOD2","TCP1","DNAJB6","RHOB","H3F3B","MXD1","AC020916.1","IER5","JUNB","TNFSF14")
IGgenes = grep("^IG[KLH]", all_genes, value=T)
HLAgenes = grep("^HLA-", all_genes, value=T)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
cell_cycle_genes = c(s.genes,g2m.genes)
genes_mask_clust = c(cell_cycle_genes,HB_genes, metallothionein, rpgenes,other,IGgenes,HLAgenes,MTRN_genes,mtgenes)
genes_mask_plot = c(HB_genes, metallothionein, rpgenes,other,IGgenes,MTRN_genes,mtgenes)
```


## 2.2.  DimReduction
```{r }
DefaultAssay(atlas_skin_myelo) = "RNA"
atlas_skin_myelo = NormalizeData(atlas_skin_myelo) 
atlas_skin_myelo <- CellCycleScoring(atlas_skin_myelo, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
atlas_skin_myelo = FindVariableFeatures(atlas_skin_myelo, nfeatures = 2000)
vargenes = setdiff(VariableFeatures(atlas_skin_myelo), genes_mask_clust)
VariableFeatures(atlas_skin_myelo)= vargenes
atlas_skin_myelo <- ScaleData(atlas_skin_myelo, assay = "RNA", features = vargenes, vars.to.regress =  c("S.Score", "G2M.Score"))
atlas_skin_myelo <- RunPCA(atlas_skin_myelo,verbose = FALSE, npcs = 50, reduction.name = "pca", features = vargenes)
nPC= 19
ElbowPlot(atlas_skin_myelo, ndims = 50, reduction = "pca")+geom_vline(xintercept=nPC,  color="red")

```

## 2.3. Clustering

```{r fig.height=9, fig.width=12}
# 33952 features across 24700
atlas_skin_myelo <- FindNeighbors(atlas_skin_myelo, reduction = "pca", dims = 1:nPC)
atlas_skin_myelo <- FindClusters(atlas_skin_myelo, resolution = 0.3)
atlas_skin_myelo <- RunUMAP(atlas_skin_myelo, reduction = "pca",min.dist = 0.5,spread = 1,  dims = 1:nPC, reduction.name = "umap")
DimPlot(atlas_skin_myelo)
atlas_skin_myelo <- FindClusters(atlas_skin_myelo, resolution = 0.2)

# Cluster 6 are DC, subset object to cluster 6 and recluster
cells= Cells(atlas_skin_myelo)
FeaturePlot(atlas_skin_myelo, features=c("FCER1A","CD1C","C5AR1","FOLR2","CD207","CLEC9A","XCR1","CLEC10A"), cells = cells, ncol=4, reduction= "umap") & NoAxes()&coord_equal()
markers = presto::wilcoxauc(atlas_skin_myelo,group.by = "seurat_clusters")

markers_dist = markers %>% filter(!feature %in% genes_mask_clust) %>% arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>%  group_by(group) %>% arrange(desc(auc)) %>% slice_head(n = 30)

DimPlot(atlas_skin_myelo, group.by = "seurat_clusters")

```

## 2.4. Visualization and removing junk clusters
Heatmaps
```{r , fig.height=11, fig.width=12.5, eval=T}
atlas_skin_myelo@active.ident = factor(atlas_skin_myelo$seurat_clusters)
genes2show= c("CD1C","FCER1A","CLEC10A","CD207","CD1A","CCR6","EPCAM")
markers = presto::wilcoxauc(atlas_skin_myelo, group_by="seurat_clusters")
top_markers= presto_topn_auc(markers,object= atlas_skin_myelo@active.ident,min_pct_in=10, genes_filter=genes_mask_clust, top_n = 12,minavgExpr=0.1)

atlas_skin_myelo <- ScaleData(atlas_skin_myelo,features=c(genes2show,top_markers$feature))
atlas_skin_myelo$log_nCount = log10(atlas_skin_myelo$nCount_RNA)

plot_heatmap2(atlas_skin_myelo, markers =c(genes2show,top_markers$feature),sort_var = c("seurat_clusters","bodysite","log_nCount"),
              anno_var = c("seurat_clusters","bodysite","log_nCount"),
              anno_colors = list(pal,pal,dim_pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),
              row_font_size = 7)
VlnPlot(atlas_skin_myelo, "log_nCount")+plot_layout(nrow=3, ncol=2)

# remove clusters: 2 KRT epithelial+ Collagenes, 6, 9 & 11 low quality, 8 B cells, 10 collagenes 
plot(density(atlas_skin_myelo$log_nCount))
atlas_skin_myelo = subset(atlas_skin_myelo, idents= c(2,6,9,11,8,10), invert=T)
# 34621 features across 8975
```
## 2.5. Repeating DimReduction
```{r}
atlas_skin_myelo = FindVariableFeatures(atlas_skin_myelo, nfeatures = 2000)
vargenes = setdiff(VariableFeatures(atlas_skin_myelo), genes_mask_clust)
VariableFeatures(atlas_skin_myelo)= vargenes
atlas_skin_myelo <- ScaleData(atlas_skin_myelo, assay = "RNA", features = vargenes, vars.to.regress =  c("S.Score", "G2M.Score"))
atlas_skin_myelo <- RunPCA(atlas_skin_myelo,verbose = FALSE, npcs = 50, reduction.name = "pca", features = vargenes)
nPC= 17
ElbowPlot(atlas_skin_myelo, ndims = 50, reduction = "pca")+geom_vline(xintercept=nPC,  color="red")+plot_layout(nrow=2, ncol=2)
```

## 2.6. Repeating Clustering

```{r fig.height=9, fig.width=12}
# 33952 features across 24700
atlas_skin_myelo <- FindNeighbors(atlas_skin_myelo, reduction = "pca", dims = 1:nPC)
atlas_skin_myelo <- FindClusters(atlas_skin_myelo, resolution = 0.3)
atlas_skin_myelo <- RunUMAP(atlas_skin_myelo, reduction = "pca",min.dist = 0.5,spread = 1,  dims = 1:nPC, reduction.name = "umap")
DimPlot(atlas_skin_myelo)
```

## 2.6. Annotation and reordering of clusters
```{r eval=F}
# p <- SCpubr::do_CorrelationPlot(sample = atlas_skin_myelo,
#                                 group.by = "RNA_snn_res.0.3",
#                                 legend.position = "right")
# levels(p$data$x)
atlas_skin_myelo$seurat_clusters = factor(atlas_skin_myelo$RNA_snn_res.0.3, levels= c("4","7","3","0","8","5","6","2","1"))  # levels(p$data$x)

atlas_skin_myelo$res.0.3 = plyr::mapvalues(atlas_skin_myelo$seurat_clusters, from = c("4","7","3","0","8","5","6","2","1"), to =c(1:9))
atlas_skin_myelo$annot_lev3 = plyr::mapvalues(atlas_skin_myelo$res.0.3, from = c(1:9), to =c("DC1","mature_DC","DC2","DC2_CCR7","DC2_cyc","DC3","DC_other","DC2_CD207","MonoMac"))
atlas_skin_myelo@active.ident = atlas_skin_myelo$annot_lev3
```

## 2.7 Save final processed object
```{r eval=T}
# qsave(atlas_skin_myelo, "../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3_myelo_reprocessed.qs", nthreads = 8)
atlas_skin_myelo = qread("../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3_myelo_reprocessed.qs", nthreads = 8)
```

```{r eval=T}
atlas_skin_myelo = qread("../../../../../data_global/public/single_cellRNA/Skin/Atlas_multi_skin/seurat_v3_myelo_reprocessed.qs", nthreads = 8)
DimPlot(atlas_skin_myelo, group.by = c("RNA_snn_res.0.3", "res.0.3", "annot_lev3"), ncol=3, reduction= "umap") & NoAxes()&coord_equal()
```

## 2.8. Visualization
Heatmaps
```{r , fig.height=11, fig.width=12.5, eval=T}
atlas_skin_myelo@active.ident = factor(atlas_skin_myelo$annot_lev3)
genes2show= c("CD1C","FCER1A","CLEC10A","CD207","CD1A","ITGAE","IL18R1","CCR6","EPCAM","LAMP3","IDO1","CCR7","CLEC9A")
markers = presto::wilcoxauc(atlas_skin_myelo, group_by="annot_lev3")
top_markers= presto_topn_auc(markers,object= atlas_skin_myelo@active.ident,min_pct_in=10, genes_filter=genes_mask_plot, top_n = 12,minavgExpr=0.1)

atlas_skin_myelo <- ScaleData(atlas_skin_myelo,features=c(genes2show,top_markers$feature))

plot_heatmap2(atlas_skin_myelo, markers =c(genes2show,top_markers$feature),sort_var = c("annot_lev3","bodysite"),
              anno_var = c("annot_lev3","bodysite"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),
              row_font_size = 7)

top_markers= presto_topn_auc(markers,object= atlas_skin_myelo@active.ident,min_pct_in=10, genes_filter=genes_mask_plot, top_n = 120,minavgExpr=0.1)

atlas_skin_myelo <- ScaleData(atlas_skin_myelo,features=c(top_markers$feature))

plot_heatmap2(atlas_skin_myelo, markers =c(top_markers$feature[top_markers$group=="DC2_CD207"][1:100]),sort_var = c("annot_lev3","bodysite"),
              anno_var = c("annot_lev3","bodysite"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),padding=unit(c(0, 0, 0, 6.5), "cm"), 
              row_font_size = 7)

VlnPlot(atlas_skin_myelo, "log_nCount")+plot_layout(nrow=3, ncol=2)

# remove clusters: 2 KRT epithelial+ Collagenes, 6, 9 & 11 low quality, 8 B cells, 10 collagenes 
plot(density(atlas_skin_myelo$log_nCount))
atlas_skin_myelo = subset(atlas_skin_myelo, idents= c(2,6,9,11,8,10), invert=T)
# 34621 features across 8975

atlas_skin_myelo
VlnPlot(atlas_skin_myelo, features =c("CD1C","FCER1A","CLEC10A","CD207","CD1A","EPCAM","CCR6","ITGAE","HPGDS","IL18R1","SIGLEC9","CD72","LAMP3","IDO1","CCR7","CLEC9A","CD14"), fill="ident",stack=T, flip=T)+NoLegend()+plot_layout(ncol=3)

markers = presto::wilcoxauc(atlas_skin_myelo, group_by="annot_lev3")
library(magrittr)

markers = presto::wilcoxauc(atlas_skin_myelo) 

markers$feature = UpdateSymbolsHuman(markers$feature)

top_markers = markers %>% filter(!feature %in% genes_mask_plot & avgExpr>0.1 & logFC>0.1  & pct_in>10) %>% group_by(group) %>% arrange(desc(auc),.by_group =T )

top_markers_distinct = markers %>% filter(!feature %in% genes_mask_plot &  avgExpr>0.1 & pct_in>10 & logFC>0.1)  %>% arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>% group_by(group) %>% arrange(desc(auc), .by_group =T)

WriteXLS::WriteXLS(top_markers, "results/multisite_skin_atlas_myeloid_clusters_markers.xlsx")
WriteXLS::WriteXLS(top_markers_distinct, "results/multisite_skin_atlas_myeloid_clusters_markers_distinct.xlsx")

top_markers_distinct_multi_site = readxl::read_excel("results/multisite_skin_atlas_myeloid_clusters_markers_distinct.xlsx")
top_markers_distinct_woundHealing = readxl::read_excel("results/skin_wound_healing_myeloid_clusters_markers_distinct.xlsx")

share_signature_100= intersect(top_markers_distinct_multi_site$feature[top_markers_distinct_multi_site$group=="DC2_CD207"], top_markers_distinct_woundHealing$feature[top_markers_distinct_woundHealing$group=="DC2_CD207"])[1:100]

atlas_skin_myelo <- ScaleData(atlas_skin_myelo,features=c(share_signature_100))

plot_heatmap2(atlas_skin_myelo, markers =share_signature_100,sort_var = c("annot_lev3","bodysite"),
              anno_var = c("annot_lev3","bodysite"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),padding=unit(c(0, 0, 0, 10), "cm"), 
              row_font_size = 7)

seu_sub <- ScaleData(seu_sub,features=c(share_signature_100))
seu_sub$seurat_clusters

plot_heatmap2(seu_sub, markers =share_signature_100,sort_var = c("annot_lev3","timepoint"),
              anno_var = c("annot_lev3","timepoint"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),padding=unit(c(0, 0, 0, 10), "cm"), 
              row_font_size = 7)
```


 ###########################################################################
 ########################################################################### 
 ########################################################################### 

# 3. Bcc_and_normal Spatial-skin-atlas
```{r Load Data, eval=F}
# data was downloaded from
library("Seurat")
library("anndata")
# https://github.com/mojaveazure/seurat-disk/issues/109
print("Convert from Scanpy to Seurat...")
data <- read_h5ad("../../../../../data_global/public/single_cellRNA/Skin/Spatial-skin-atlas/bcc_and_normal-CG_portal_fat.h5ad")
str(data$raw$X)
counts= data$raw$X
colnames(counts) = colnames(data$X)
rownames(counts) = rownames(data$X)

seurat_obj <- CreateSeuratObject(counts = t(counts), meta.data = data$obs)
str(data$obsm$X_umap)
umap= data$obsm$X_umap
rownames(umap) = colnames(seurat_obj)
umap = CreateDimReducObject(umap, key="umap")
seurat_obj@reductions$umap = umap
# 32983 features across 155401
# qs::qsave(seurat_obj, "../../../../../data_global/public/single_cellRNA/Skin/Spatial-skin-atlas/bcc_and_normal-CG_portal_fat_seurat.qs", nthreads = 8)
```

```{r}
seurat_obj= qs::qread("../../../../../data_global/public/single_cellRNA/Skin/Spatial-skin-atlas/bcc_and_normal-CG_portal_fat_seurat.qs")
unique(seurat_obj$`04_celltypes`)

DimPlot(seurat_obj, group.by = "04_celltypes")
seurat_sub = subset(seurat_obj, `04_celltypes` %in% c("Monocytes","Macrophages","DC"))
# 23602 cells
```

```{r Prepare gene masks, eval=T}
all_genes <- rownames(seurat_sub)
HB_genes <- c("HBB","HBA1","HBA2", "HBD","HBE1","HBG1","HBM","HBBP1","HBQ1","AHSP")
metallothionein=c('MT1HL1','MT2A','MT1E','MT1M','MT1A','MT1B','MT1F','MT1G','MT1H','MT1X')
rpgenes= grep("^RP[LS0-9]", all_genes, value=T) # 
other=c("MALAT1","XIST","JCHAIN")
mtgenes=grep("^MT-", all_genes, value=T) 
MTRN_genes=grep("^MTRN", all_genes, value=T)
# stress_genes = c("HSPA1B","HSPA6","DNAJB1","HSPA1A","HSPB1","JUN","UBB","DNAJB4","BAG3","UBC","SOD1","RGS2","EIF4A2","SERPINH1","DEDD2","CLK1","AHSA1","COMMD6","HSP90AA1","HSPH1","HSPE1","HSPD1","HSP90AB1","DUSP1","HSPA8","FOS","DNAJA1","PPP1R15A","CACYBP","KLF6","ZFAND2A","NR4A1","FOSB","DNAJA4","CHORDC1","MRPL18","BTG2","FKBP4","NEU1","ATF3","MYADM","ERN1","DOK2","GADD45B","TXNIP","NABP1","SOD2","TCP1","DNAJB6","RHOB","H3F3B","MXD1","AC020916.1","IER5","JUNB","TNFSF14")
IGgenes = grep("^IG[KLH]", all_genes, value=T)
HLAgenes = grep("^HLA-", all_genes, value=T)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
cell_cycle_genes = c(s.genes,g2m.genes)
genes_mask_clust = c(cell_cycle_genes,HB_genes, metallothionein, rpgenes,other,IGgenes,HLAgenes,MTRN_genes,mtgenes)
genes_mask_plot = c(HB_genes, metallothionein, rpgenes,other,IGgenes,MTRN_genes,mtgenes)
```

## 3.1 DimReduction
```{r }
DefaultAssay(seurat_sub) = "RNA"
seurat_sub = NormalizeData(seurat_sub)
seurat_sub <- CellCycleScoring(seurat_sub, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
seurat_sub = FindVariableFeatures(seurat_sub)
vargenes = setdiff(VariableFeatures(seurat_sub), genes_mask_clust)
seurat_sub <- ScaleData(seurat_sub, assay = "RNA", features = vargenes, vars.to.regress =  c("S.Score", "G2M.Score"))
seurat_sub <- RunPCA(seurat_sub,verbose = FALSE, npcs = 50, reduction.name = "pca", features = vargenes)
nPC= 18
ElbowPlot(seurat_sub, ndims = 50, reduction = "pca")+geom_vline(xintercept=nPC,  color="red")+ plot_layout(nrow=2, ncol=2)
```

## 3.2. Clustering

```{r fig.height=9, fig.width=12}
seurat_sub <- RunUMAP(seurat_sub, reduction = "pca",min.dist = 0.5,spread = 1,  dims = 1:nPC, reduction.name = "umap")
DimPlot(seurat_sub, group.by = "04_celltypes")

# 32983 features across 23602
seurat_sub <- FindNeighbors(seurat_sub, reduction = "pca", dims = 1:nPC)
seurat_sub <- FindClusters(seurat_sub, resolution = 0.35)
# Cluster 6 are DC, subset object to cluster 6 and recluster
DimPlot(seurat_sub)
 
cells= Cells(seurat_sub)
FeaturePlot(seurat_sub, features=c("FCER1A","CD1C","C5AR1","FOLR2","CD207","SPP1","SELENOP","TREM2"), cells = cells, ncol=4, reduction= "umap") & NoAxes()&coord_equal()
markers = presto::wilcoxauc(seurat_sub)
markers_dist = markers %>% filter(!feature %in% genes_mask_clust) %>% arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>%  group_by(group) %>% arrange(desc(auc)) %>% slice_head(n = 10)
# plan("multicore", workers = 8) # unstable in rstudio
```
Heatmaps
```{r , fig.height=10.5, fig.width=12, eval=T}
seu_small= subset(seurat_sub, dowsample=500, group.by="seurat_clusters")

seu_small@active.ident = seu_small$seurat_clusters
markers = presto::wilcoxauc(seu_small)
top_markers= presto_topn_auc(markers,object= seu_small@active.ident,min_pct_in=10, genes_filter=genes_mask_plot, top_n = 10,minavgExpr=0.1)

seu_small <- ScaleData(seu_small,features=c(top_markers$feature))

plot_heatmap2(seu_small, markers =unique(c(top_markers$feature)),sort_var = c("seurat_clusters","05_subcelltypes","02_group"),
              anno_var = c("seurat_clusters","05_subcelltypes","02_group"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,T,T,T,F),
              row_font_size = 7)


seu_small@active.ident = seu_small$`05_subcelltypes`

markers = presto::wilcoxauc(seu_small, group_by="05_subcelltypes")

top_markers= presto_topn_auc(markers,object= seu_small@active.ident,min_pct_in=10, genes_filter=genes_mask_plot, top_n = 80,minavgExpr=0.1)

seu_small <- ScaleData(seu_small,features=c(top_markers$feature))

plot_heatmap2(seu_small, markers =unique(c(top_markers$feature)),sort_var = c("05_subcelltypes","seurat_clusters","02_group"),
              anno_var = c("05_subcelltypes","seurat_clusters","02_group"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,T,T,T,F),
              row_font_size = 7)

top_markers= presto_topn_auc(markers,object= seu_small@active.ident,min_pct_in=10, genes_filter=genes_mask_plot, top_n = 100,minavgExpr=0.1)

seu_small <- ScaleData(seu_small,features=c(top_markers$feature))

plot_heatmap2(seu_small, markers =unique(c(top_markers$feature[top_markers$group=="12"])),sort_var = c("seurat_clusters","05_subcelltypes","02_group"),
              anno_var = c("seurat_clusters","05_subcelltypes","02_group"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,T,T,T,F),
              row_font_size = 7)

markers %>% filter(!feature %in% genes_mask_clust)  %>%  arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>% group_by(group) %>% arrange(desc(auc)) %>% filter(group=="9") %>% slice_head(n = 200) %>% pull(feature)

```

 ###########################################################################
 ###########################################################################
 ###########################################################################
 
 
# 4. Chronic Inf Skin disease atlas

```{r}
atlas_skin_myelo = readRDS("../../../../../data_global/public/single_cellRNA/Skin/Chronic Inf Skin disease/Manuscript Object.rds")
rownames(atlas_skin_myelo@assays$ADT)
VlnPlot(subset(atlas_skin_myelo,type=="APC") , "EPCAM", group.by = "Ident1")

##### remove mouse genes from the Objects / contamination?
mouse_genes <- grep("^mm10---", rownames(atlas_skin_myelo), value = TRUE)
human_genes = setdiff(rownames(atlas_skin_myelo), mouse_genes)
atlas_skin_myelo <- subset(atlas_skin_myelo, features = human_genes)
# 33570 features across 145810
```

Extract metadata from sample names
```{r}
all_genes <- rownames(atlas_skin_myelo)
HB_genes <- c("HBB","HBA1","HBA2", "HBD","HBE1","HBG1","HBM","HBBP1","HBQ1","AHSP")
metallothionein=c('MT1HL1','MT2A','MT1E','MT1M','MT1A','MT1B','MT1F','MT1G','MT1H','MT1X')
rpgenes= grep("^RP[LS0-9]", all_genes, value=T) # 
other=c("MALAT1","XIST","JCHAIN",grep("^IGLL", all_genes, value=T))
mtgenes=grep("^MT-", all_genes, value=T) 
MTRN_genes=grep("^MTRN", all_genes, value=T)
# stress_genes = c("HSPA1B","HSPA6","DNAJB1","HSPA1A","HSPB1","JUN","UBB","DNAJB4","BAG3","UBC","SOD1","RGS2","EIF4A2","SERPINH1","DEDD2","CLK1","AHSA1","COMMD6","HSP90AA1","HSPH1","HSPE1","HSPD1","HSP90AB1","DUSP1","HSPA8","FOS","DNAJA1","PPP1R15A","CACYBP","KLF6","ZFAND2A","NR4A1","FOSB","DNAJA4","CHORDC1","MRPL18","BTG2","FKBP4","NEU1","ATF3","MYADM","ERN1","DOK2","GADD45B","TXNIP","NABP1","SOD2","TCP1","DNAJB6","RHOB","H3F3B","MXD1","AC020916.1","IER5","JUNB","TNFSF14")
IGgenes = grep("^IG[KLH]", all_genes, value=T)
HLAgenes = grep("^HLA-", all_genes, value=T)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
cell_cycle_genes = c(s.genes,g2m.genes)
genes_mask_clust = c(cell_cycle_genes,HB_genes, metallothionein, rpgenes,other,IGgenes,HLAgenes,MTRN_genes,mtgenes)
genes_mask_plot = c(HB_genes, metallothionein, rpgenes,other,IGgenes,MTRN_genes,mtgenes)
```

## 4.1. DimReduction
```{r }
str(atlas_skin_myelo@meta.data)
tabl(atlas_skin_myelo$Ident2)
tabl(skin_sub$Ident2)
skin_sub = subset(atlas_skin_myelo, Ident2 %in% c("Mac1","Mac2","Mono","InfMono","moDC1","moDC2","moDC3","LC1","LC2","LC3","DC1","DC2","migDC","DC3","Mac2","Mac3","Mac4"), invert=F)
Idents(skin_sub)= skin_sub$Ident2
skin_sub = subset(skin_sub, downsample=500)
DefaultAssay(skin_sub) = "RNA"
skin_sub = NormalizeData(skin_sub) 
skin_sub <- CellCycleScoring(skin_sub, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
skin_sub = FindVariableFeatures(skin_sub, nfeatures = 2000)
VariableFeatures(skin_sub)= vargenes
vargenes = setdiff(VariableFeatures(skin_sub), genes_mask_clust)
skin_sub <- ScaleData(skin_sub, assay = "RNA", features = vargenes, vars.to.regress =  c("S.Score", "G2M.Score"))
skin_sub <- RunPCA(skin_sub,verbose = FALSE, npcs = 50, reduction.name = "pca", features = vargenes)
nPC= 18

ElbowPlot(skin_sub, ndims = 50, reduction = "pca")+geom_vline(xintercept=nPC,  color="red")
skin_sub <- RunUMAP(skin_sub, reduction = "pca",min.dist = 0.5,spread = 1,  dims = 1:nPC, reduction.name = "umap")
```

## 4.2. Clustering

```{r fig.height=9, fig.width=12}
# 33952 features across 24700
skin_sub$Ident2 = factor(skin_sub$Ident2 , levels=  c("Mac1","Mac2","Mac3","Mac4","Mono","InfMono","moDC1","moDC2","moDC3","LC1","LC2","LC3","DC1","DC2","migDC","DC3"))
skin_sub <- FindNeighbors(skin_sub, reduction = "pca", dims = 1:nPC)
skin_sub <- FindClusters(skin_sub, resolution = 0.5)
skin_sub <- FindClusters(skin_sub, resolution = 0.8)
skin_sub <- FindClusters(skin_sub, resolution = 1.5)
# Cluster 6 are DC, subset object to cluster 6 and recluster
cells= Cells(skin_sub)
FeaturePlot(skin_sub, features=c("FCER1A","CD1C","C5AR1","FOLR2","CD207","CLEC9A","XCR1","CLEC10A"), cells = cells, ncol=4, reduction= "umap") & NoAxes()&coord_equal()
markers = presto::wilcoxauc(skin_sub,group.by = "seurat_clusters")

markers_dist = markers %>% filter(!feature %in% genes_mask_clust) %>% arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>%  group_by(group) %>% arrange(desc(auc)) %>% slice_head(n = 30)

DimPlot(skin_sub, group.by = "seurat_clusters")

saveRDS(skin_sub, "./skin_sub.rds", compress=F)
skin_sub = readRDS("skin_sub.rds")
```

Heatmaps
```{r , fig.height=10.5, fig.width=12.5, eval=F}
skin_sub@active.ident = factor(skin_sub$seurat_clusters)
genes2show= c("CD1C","FCER1A","CLEC10A","CD207","CD1A","CCR6","EPCAM")
markers = presto::wilcoxauc(skin_sub, group_by="seurat_clusters")
top_markers= presto_topn_auc(markers,object= skin_sub@active.ident,min_pct_in=10, genes_filter=genes_mask_clust, top_n = 8,minavgExpr=0.1)

skin_sub <- ScaleData(skin_sub,features=c(genes2show,top_markers$feature))

plot_heatmap2(skin_sub, markers =c(genes2show,top_markers$feature),sort_var = c("seurat_clusters","Ident2","orig.ident"),
              anno_var = c("seurat_clusters","Ident2","orig.ident"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),
              row_font_size = 6)

skin_sub@active.ident = factor(skin_sub$Ident2)

markers = presto::wilcoxauc(skin_sub, group_by="Ident2")

top_markers= presto_topn_auc(markers,object= skin_sub@active.ident,min_pct_in=20, genes_filter=genes_mask_clust, top_n = 10,minavgExpr=0.1)

skin_sub <- ScaleData(skin_sub,features=c(top_markers$feature))

plot_heatmap2(skin_sub, markers = unique(c(top_markers$feature)),sort_var = c("Ident2","seurat_clusters","orig.ident"),
              anno_var = c("Ident2","seurat_clusters","orig.ident"),
              anno_colors = list(pal,pal,pal),
              hm_colors = c("purple","black","yellow"),
              hm_limit = c(-2,0,2),legend2show = c(T,T,F,T,T,F),
              row_font_size = 7)

markers %>% filter(!feature %in% genes_mask_clust)  %>%  arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>% group_by(group) %>% arrange(desc(auc)) %>% filter(group=="LC") %>% slice_head(n = 200) %>% pull(feature)

markers %>% filter(!feature %in% genes_mask_clust)  %>%  arrange(desc(avgExpr)) %>% distinct(feature,.keep_all = TRUE) %>% group_by(group) %>% arrange(desc(auc)) %>% filter(feature=="CD207")
top_markers %>% filter(group=="2") 

```


